---
title: "ARCH&GARCH"
author: "Lin Meng"
date: "3/29/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}
set.seed(666)
library(plotly)
library(ggfortify)
library(ggplot2)
library(fpp)
library(stats)
library(tsbox)
library(stargazer)
library(tidyverse)
library(kableExtra)
library(forecast)
library(astsa)
library(car)
library(TSA)
library(fGarch) 
library(dynlm)
```


## Load Data
```{r}
johnson <- read.csv("data/Johnson.csv ")
moderna <- read.csv("data/Moderna.csv ")
pfizer <- read.csv("data/Pfizer.csv ")
```

## johnson

### time series Plot
```{r}
johnson.ts <- ts_ts(ts_long(johnson[,c(1,5)]))
johnson.ts <- na.remove(johnson.ts)
fig1 <- autoplot(johnson.ts,colour='#029386') + labs(x="Date", y="Closed", title="Johnson&Johnson Closed Value By Date")
fig1
```

### ACF + PACF
From the above acf and pacf plot and previous analysis, the best arima model should be ARIMA(1,1,3)
```{r}
par(mfrow=c(1,2))
acf(johnson.ts)
pacf(johnson.ts)
```

### Train
```{r}
johnson.train <- johnson.ts[1:floor(length(johnson.ts)*0.8)]
johnson.test <- johnson.ts[(ceiling(length(johnson.ts)*0.8)+1):length(johnson.ts)]
johnson.ar <- arima(johnson.train, order=c(1,1,3))
sqrt(mean(johnson.ar$residuals^2))
```

```{r}
sarima(johnson.ts,1,1,3)
```

### Save Residuels 
From the below acf plot, both p and q can be 1 or2. 
```{r}
par(mfrow=c(1,2))
acf(johnson.ar$residuals)
pacf(johnson.ar$residuals)
```

Comparing the RMSE, the best model is ARMA(1,1,3) + Garch(2,1)
```{r}
garch1 <- garch(johnson.ar$residuals,order=c(1,1),trace=F)
garch2 <- garch(johnson.ar$residuals,order=c(1,2),trace=F)
garch3 <- garch(johnson.ar$residuals,order=c(2,1),trace=F)
garch4 <- garch(johnson.ar$residuals,order=c(2,2),trace=F)
g1 <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(g1) <- c('Model','RMSE')
g1[1,]<- c('GARCH(1,1)',sqrt(mean(na.omit(garch1$residuals^2))))
g1[2,]<- c('GARCH(1,2)',sqrt(mean(na.omit(garch2$residuals^2))))
g1[3,]<- c('GARCH(2,1)',sqrt(mean(na.omit(garch3$residuals^2))))
g1[4,]<- c('GARCH(2,2)',sqrt(mean(na.omit(garch4$residuals^2))))
g1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

The p-value of Box Ljung test is smaller than 0.05. It means there is significant correlations in residuals.
```{r}
summary(garch3)

```
## pfizer

### time series Plot
```{r}
pfizer.ts <- ts_ts(ts_long(pfizer[,c(1,5)]))
pfizer.ts <- na.remove(pfizer.ts)
fig1 <- autoplot(pfizer.ts,colour='#e6daa6') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig1
```

### ACF + PACF
From the above acf and pacf plot and previous analysis, the best arima model should be ARIMA(1,1,3)
```{r}
par(mfrow=c(1,2))
acf(pfizer.ts)
pacf(pfizer.ts)
```

### Train
```{r}
pfizer.train <- pfizer.ts[1:floor(length(pfizer.ts)*0.8)]
pfizer.test <- pfizer.ts[(ceiling(length(pfizer.ts)*0.8)+1):length(pfizer.ts)]
pfizer.ar <- arima(pfizer.train, order=c(1,1,3))
sqrt(mean(pfizer.ar$residuals^2))
```

```{r}
sarima(pfizer.ts,1,1,3)
```

### Save Residuels 
From the below acf plot, both p and q can be 1 or2. 
```{r}
par(mfrow=c(1,2))
acf(pfizer.ar$residuals)
pacf(pfizer.ar$residuals)
```

Comparing the RMSE, the best model is ARMA(1,1,3) + Garch(1,2)
```{r}
garch1 <- garch(pfizer.ar$residuals,order=c(1,1),trace=F)
garch2 <- garch(pfizer.ar$residuals,order=c(1,2),trace=F)
garch3 <- garch(pfizer.ar$residuals,order=c(2,1),trace=F)
garch4 <- garch(pfizer.ar$residuals,order=c(2,2),trace=F)
g1 <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(g1) <- c('Model','RMSE')
g1[1,]<- c('GARCH(1,1)',sqrt(mean(na.omit(garch1$residuals^2))))
g1[2,]<- c('GARCH(1,2)',sqrt(mean(na.omit(garch2$residuals^2))))
g1[3,]<- c('GARCH(2,1)',sqrt(mean(na.omit(garch3$residuals^2))))
g1[4,]<- c('GARCH(2,2)',sqrt(mean(na.omit(garch4$residuals^2))))
g1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

The p-value of Box Ljung test is smaller than 0.05. It means quite significant correlations in residuals.
```{r}
summary(garch2)

```

## moderna

### time series Plot
```{r}
moderna.ts <- ts_ts(ts_long(moderna[,c(1,5)]))
moderna.ts <- na.remove(moderna.ts)
fig1 <- autoplot(moderna.ts,colour='#ff796c') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig1
```

### ACF + PACF
From the above acf and pacf plot and previous analysis, the best arima model should be ARIMA(1,1,3)
```{r}
par(mfrow=c(1,2))
acf(moderna.ts)
pacf(moderna.ts)
```

### Train
```{r}
moderna.train <- moderna.ts[1:floor(length(moderna.ts)*0.8)]
moderna.test <- moderna.ts[(ceiling(length(moderna.ts)*0.8)+1):length(moderna.ts)]
moderna.ar <- arima(moderna.train, order=c(3,1,3))
sqrt(mean(moderna.ar$residuals^2))
```

```{r}
sarima(moderna.ts,3,1,3)
```

### Save Residuels 
From the below acf plot, both p and q can be 1 or2. 
```{r}
par(mfrow=c(1,2))
acf(moderna.ar$residuals)
pacf(moderna.ar$residuals)
```

Comparing the RMSE, the best model is ARMA(3,1,3) + Garch(1,2)
```{r}
garch1 <- garch(moderna.ar$residuals,order=c(1,1),trace=F)
garch2 <- garch(moderna.ar$residuals,order=c(1,2),trace=F)
garch3 <- garch(moderna.ar$residuals,order=c(2,1),trace=F)
garch4 <- garch(moderna.ar$residuals,order=c(2,2),trace=F)
g1 <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(g1) <- c('Model','RMSE')
g1[1,]<- c('GARCH(1,1)',sqrt(mean(na.omit(garch1$residuals^2))))
g1[2,]<- c('GARCH(1,2)',sqrt(mean(na.omit(garch2$residuals^2))))
g1[3,]<- c('GARCH(2,1)',sqrt(mean(na.omit(garch3$residuals^2))))
g1[4,]<- c('GARCH(2,2)',sqrt(mean(na.omit(garch4$residuals^2))))
g1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

The p-value of Box Ljung test is smaller than 0.05. It means quite significant correlations in residuals.
```{r}
summary(garch2)

```