---
title: "ARIMA"
author: "Lin Meng"
date: "2022/2/25"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(ggfortify)
library(ggplot2)
library(fpp)
library(stats)
library(tsbox)
library(stargazer)
library(tidyverse)
library(kableExtra)
library(forecast)
library(astsa)
```

## Load Data
```{r}
johnson <- read.csv("data/Johnson.csv ")
moderna <- read.csv("data/Moderna.csv ")
pfizer <- read.csv("data/Pfizer.csv ")
vaccine <- read.csv("data/vaccine.csv")
```

## Johnson

### 1 & 2
#### Original ACF Plot 
```{r}
johnson.ts <- ts_ts(ts_long(johnson[,c(1,5)]))
johnson.ts<-na.remove(johnson.ts)
acf(johnson.ts)
```
#### b ADF Test
```{r}
adf.test(johnson.ts)
```
```{r}
acf(log(johnson.ts))
adf.test(log(johnson.ts))
```

### 3
#### a Detrend Data
```{r}
johnson.df<-diff(johnson.ts)
```

#### b Plot the differenced data and comment.
```{r}
fig1 <- autoplot(johnson.ts,series = 'Origin') + 
  autolayer(johnson.df,series = 'Difference')
fig1 <- fig1 +labs(title = "Johnson&Johnson Origin and Difference",
                   x='Date',y='Value')
fig1
```

#### c ADF Test
```{r}
acf(johnson.df)
adf.test(johnson.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(johnson.df)
pacf(johnson.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(johnson.ts, order=c(1, 1, 1),include.drift = TRUE)
fit2 <- Arima(johnson.ts, order=c(1, 1, 3),include.drift = TRUE)
fit3 <- Arima(johnson.ts, order=c(3, 1, 1),include.drift = TRUE)
fit4 <- Arima(johnson.ts, order=c(3, 1, 3),include.drift = TRUE)
r1 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r1) <- c('Model','AIC','BIC','AICC')
r1[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r1[2,]<- c('ARIMA(1,1,3)',fit2$aic,fit2$bic,fit2$aicc)
r1[3,]<- c('ARIMA(3,1,1)',fit3$aic,fit3$bic,fit3$aicc)
r1[4,]<- c('ARIMA(3,1,3)',fit4$aic,fit4$bic,fit4$aicc)
r1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
johnson.train <- johnson.ts[1:floor(length(johnson.ts)*0.8)]
johnson.test <- johnson.ts[(ceiling(length(johnson.ts)*0.8)+1):length(johnson.ts)]
johnson.ar <- arima(johnson.train, order=c(1,1,3))
summary(johnson.ar)
```
$$  X_{t} =  1.5264)X_{t-1}-0.5264X_{t-2}+w_{t}-0.6345w_{t-1}+1.303w_{t-2}-0.0905w_{t-3}$$
### 6
#### full model diagnostics
```{r}
sarima(johnson.train,1,1,1)
```

### 7 
#### auto.arima
```{r}
auto.arima(johnson.train)
```
### 8
#### Forecast
```{r}
pred1 <- forecast(johnson.ar,length(johnson.test)+50)
autoplot(pred1) 
autoplot(johnson.ts)
```


### Sarima(p,d,q) x(P,D,Q)
```{r}
safit1 <- sarima(johnson.train,1,1,3,0,1,1,1)
safit2 <- sarima(johnson.train,1,1,3,0,1,2,1)
safit3 <- sarima(johnson.train,1,1,3,0,2,1,1)
safit4 <- sarima(johnson.train,1,1,3,0,2,2,1)
sa1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(sa1) <- c('Model','AIC','BIC')
sa1[1,]<- c('SARIMA(1,1,3)x(0,1,1)',safit1$AIC,safit1$BIC)
sa1[2,]<- c('SARIMA(1,1,3)x(0,1,2)',safit2$AIC,safit2$BIC)
sa1[3,]<- c('SARIMA(1,1,3)x(0,2,1)',safit3$AIC,safit3$BIC)
sa1[4,]<- c('SARIMA(1,1,3)x(0,2,2)',safit4$AIC,safit4$BIC)
sa1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")

johnson.sar <- Arima(johnson.train,order=c(1,1,1),seasonal=c(0,1,2))
```

### Sarima Forecast
```{r}
pred2 <- forecast(johnson.sar,length(johnson.test)+50)
autoplot(pred2) 
autoplot(johnson.ts)
```

### 9 
#### Benchmark
```{r}
f1 <- meanf(johnson.train,length(johnson.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(johnson.train,length(johnson.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(johnson.train,length(johnson.test)+50)
accuracy(f3)
checkresiduals(f3)
```

```{r message=FALSE, warning=FALSE}
library(Metrics)
a1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a1) <- c('Method','MAE','MSE')
a1[1,] <- c('Mean',mae(as.numeric(f1$mean),johnson.test),
           mse(as.numeric(f1$mean),johnson.test))
a1[2,] <- c('Naive',mae(as.numeric(f2$mean),johnson.test),
           mse(as.numeric(f2$mean),johnson.test))
a1[3,] <- c('Drift',mae(as.numeric(f3$mean),johnson.test),
           mse(as.numeric(f3$mean),johnson.test))
a1[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),johnson.test),
           mse(as.numeric(pred1$mean),johnson.test))
a1[5,] <- c('SARIMA',mae(as.numeric(pred2$mean),johnson.test),
           mse(as.numeric(pred2$mean),johnson.test))
a1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
johnson.train <- ts(johnson.train)
fit2 <- Arima(johnson.train,order=c(1,1,3),include.drift=TRUE)
johnson.sar <- Arima(johnson.train,order=c(1,1,3),seasonal=c(0,1,1))
autoplot(johnson.train,series='Data') +
  autolayer(meanf(johnson.train,150),PI=FALSE,series='Mean') +
  autolayer(naive(johnson.train,150),PI=FALSE,series='Naive') +
  autolayer(rwf(johnson.train,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  autolayer(forecast(johnson.sar,150),PI=FALSE, series='SAR') +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Benchmark Stock Forecast")
```



## Pfizer

### 1 & 2
#### Original ACF Plot 
```{r}
pfizer.ts <- ts_ts(ts_long(pfizer[,c(1,5)]))
pfizer.ts<-na.remove(pfizer.ts)
acf(pfizer.ts)
```
#### b ADF Test
```{r}
adf.test(pfizer.ts)
```
```{r}
acf(log(pfizer.ts))
adf.test(log(pfizer.ts))
```


### 3
#### a Detrend Data
```{r}
pfizer.df<-diff(pfizer.ts)
```

#### b Plot the differenced data and comment.
```{r}
fig2 <- autoplot(pfizer.ts) + 
  autolayer(pfizer.df,series = 'Difference')
fig2 <- fig2 +labs(title = "Pfizer Origin and Difference",
                   x='Date',y='Value')
fig2
```

#### c ADF Test
```{r}
acf(pfizer.df)
adf.test(pfizer.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(pfizer.df)
pacf(pfizer.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(pfizer.ts, order=c(1, 1, 1),include.drift = TRUE)
fit2 <- Arima(pfizer.ts, order=c(1, 1, 3),include.drift = TRUE)
r2 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r2) <- c('Model','AIC','BIC','AICC')
r2[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r2[2,]<- c('ARIMA(1,1,3)',fit2$aic,fit2$bic,fit2$aicc)
r2 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
pfizer.train <- pfizer.ts[1:floor(length(pfizer.ts)*0.8)]
pfizer.test <- pfizer.ts[(ceiling(length(pfizer.ts)*0.8)+1):length(pfizer.ts)]
pfizer.ar <- arima(pfizer.train, order=c(1,1,3))
summary(pfizer.ar)
```
$$  X_{t} =  0.1491X_{t-1}+0.8509X_{t-2}+w_{t}+0.7819w_{t-1}-0.0198w_{t-2}+0.0927w_{t-3}$$
### 6
#### full model diagnostics
```{r}
sarima(pfizer.train,1,1,3)
```

### 7 
#### auto.arima
```{r}
auto.arima(pfizer.train)
```

### 8
#### Forecast
```{r}
pred1 <- forecast(pfizer.ar,length(pfizer.test)+50)
autoplot(pred1) 
autoplot(pfizer.ts)
```


### Sarima(p,d,q) x(P,D,Q)
```{r}
safit1 <- sarima(pfizer.train,1,1,3,0,1,1,4)
safit2 <- sarima(pfizer.train,1,1,3,0,1,2,4)
safit3 <- sarima(pfizer.train,1,1,3,0,2,1,4)
safit4 <- sarima(pfizer.train,1,1,3,0,2,2,4)
sa1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(sa1) <- c('Model','AIC','BIC')
sa1[1,]<- c('SARIMA(1,1,3)x(0,1,1)',safit1$AIC,safit1$BIC)
sa1[2,]<- c('SARIMA(1,1,3)x(0,1,2)',safit2$AIC,safit2$BIC)
sa1[3,]<- c('SARIMA(1,1,3)x(0,2,1)',safit3$AIC,safit3$BIC)
sa1[4,]<- c('SARIMA(1,1,3)x(0,2,2)',safit4$AIC,safit4$BIC)
sa1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")

pfizer.sar <- Arima(pfizer.train,order=c(1,1,1),seasonal=c(0,1,1))

```

### Sarima Forecast
```{r}
pred2 <- forecast(pfizer.sar,length(johnson.test)+50)
autoplot(pred2) 
autoplot(johnson.ts)
```

### 9 
#### Benchmark
```{r}
f1 <- meanf(pfizer.train,length(pfizer.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(pfizer.train,length(pfizer.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(pfizer.train,length(pfizer.test)+50)
accuracy(f3)
checkresiduals(f3)
```


```{r message=FALSE, warning=FALSE}
library(Metrics)
a1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a1) <- c('Method','MAE','MSE')
a1[1,] <- c('Mean',mae(as.numeric(f1$mean),pfizer.test),
            mse(as.numeric(f1$mean),pfizer.test))
a1[2,] <- c('Naive',mae(as.numeric(f2$mean),pfizer.test),
            mse(as.numeric(f2$mean),pfizer.test))
a1[3,] <- c('Drift',mae(as.numeric(f3$mean),pfizer.test),
            mse(as.numeric(f3$mean),pfizer.test))
a1[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),pfizer.test),
            mse(as.numeric(pred1$mean),pfizer.test))
a1[5,] <- c('SARIMA',mae(as.numeric(pred2$mean),pfizer.test),
            mse(as.numeric(pred2$mean),pfizer.test))
a1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
pfizer.train <- ts(pfizer.train)
fit2 <- Arima(pfizer.train,order=c(1,1,3),include.drift=TRUE)
pfizer.sar <- Arima(pfizer.train,order=c(1,1,3),seasonal=c(0,1,1))
autoplot(pfizer.train,series='Data') +
  autolayer(meanf(pfizer.train,150),PI=FALSE,series='Mean') +
  autolayer(naive(pfizer.train,150),PI=FALSE,series='Naive') +
  autolayer(rwf(pfizer.train,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  autolayer(forecast(pfizer.sar,150),PI=FALSE, series='SAR') +
  xlab("Date") + ylab("Value") +
  ggtitle("pfizer Benchmark Stock Forecast")
```

## moderna

### 1 & 2
#### Original ACF Plot 
```{r}
moderna.ts <- ts_ts(ts_long(moderna[,c(1,5)]))
moderna.ts<-na.remove(moderna.ts)
acf(moderna.ts)
```


#### b ADF Test
```{r}
adf.test(moderna.ts)
```
```{r}
acf(log(moderna.ts))
adf.test(log(moderna.ts))
```

### 3
#### a Detrend Data
```{r}
moderna.df<-diff(moderna.ts)
```

#### b Plot the differenced data and comment.
```{r}
fig1 <- autoplot(moderna.ts,series = 'Origin') + 
  autolayer(moderna.df,series = 'Difference')
fig1 <- fig1 +labs(title = "Moderna Origin and Difference",
                   x='Date',y='Value')
fig1
```

#### c ADF Test
```{r}
acf(moderna.df)
adf.test(moderna.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(moderna.df)
pacf(moderna.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(moderna.ts, order=c(2, 1, 2),include.drift = TRUE)
fit2 <- Arima(moderna.ts, order=c(2, 1, 3),include.drift = TRUE)
fit3 <- Arima(moderna.ts, order=c(3, 1, 2),include.drift = TRUE)
fit4 <- Arima(moderna.ts, order=c(3, 1, 3),include.drift = TRUE)
r3 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r3) <- c('Model','AIC','BIC','AICC')
r3[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r3[2,]<- c('ARIMA(1,1,3)',fit2$aic,fit2$bic,fit2$aicc)
r3[3,]<- c('ARIMA(3,1,1)',fit3$aic,fit3$bic,fit3$aicc)
r3[4,]<- c('ARIMA(3,1,3)',fit4$aic,fit4$bic,fit4$aicc)
r3 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
moderna.train <- moderna.ts[1:floor(length(moderna.ts)*0.8)]
moderna.test <- moderna.ts[(ceiling(length(moderna.ts)*0.8)+1):length(moderna.ts)]
moderna.ar <- arima(moderna.train, order=c(3,1,3))
summary(moderna.ar)
```
$$  X_{t} = 1.6268X_{t-1}+0.1617X_{t-2}+1.3978X_{t-3}-0.1917X_{t-4}+w_{t}-0.6345w_{t-1}+1.303w_{t-2}-0.0905w_{t-3}$$
### 6
#### full model diagnostics
```{r}
sarima(moderna.train,1,1,1)
```

### 7 
#### auto.arima
```{r}
auto.arima(moderna.train)
```

### 8
#### Forecast
```{r}
pred1 <- forecast(moderna.ar,length(moderna.test)+50)
autoplot(pred1) 
autoplot(moderna.ts)
```

### Sarima(p,d,q) x(P,D,Q)
```{r}
safit1 <- sarima(moderna.train,3,1,3,0,1,1,1)
safit2 <- sarima(moderna.train,3,1,3,0,1,2,1)
safit3 <- sarima(moderna.train,3,1,3,0,2,1,1)
safit4 <- sarima(moderna.train,3,1,3,0,2,2,1)
sa1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(sa1) <- c('Model','AIC','BIC')
sa1[1,]<- c('SARIMA(3,1,3)x(0,1,1)',safit1$AIC,safit1$BIC)
sa1[2,]<- c('SARIMA(3,1,3)x(0,1,2)',safit2$AIC,safit2$BIC)
sa1[3,]<- c('SARIMA(3,1,3)x(0,2,1)',safit3$AIC,safit3$BIC)
sa1[4,]<- c('SARIMA(3,1,3)x(0,2,2)',safit4$AIC,safit4$BIC)
sa1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")

moderna.sar <- Arima(moderna.train,order=c(3,1,3),seasonal=c(0,1,2))
```

### Sarima Forecast
```{r}
pred2 <- forecast(moderna.sar,length(moderna.test)+50)
autoplot(pred2) 
autoplot(moderna.ts)
```

### 9 
#### Benchmark
```{r}
f1 <- meanf(moderna.train,length(moderna.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(moderna.train,length(moderna.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(moderna.train,length(moderna.test)+50)
accuracy(f3)
checkresiduals(f3)
```

```{r message=FALSE, warning=FALSE}
library(Metrics)
a1 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a1) <- c('Method','MAE','MSE')
a1[1,] <- c('Mean',mae(as.numeric(f1$mean),moderna.test),
           mse(as.numeric(f1$mean),moderna.test))
a1[2,] <- c('Naive',mae(as.numeric(f2$mean),moderna.test),
           mse(as.numeric(f2$mean),moderna.test))
a1[3,] <- c('Drift',mae(as.numeric(f3$mean),moderna.test),
           mse(as.numeric(f3$mean),moderna.test))
a1[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),moderna.test),
           mse(as.numeric(pred1$mean),moderna.test))
a1[5,] <- c('SARIMA',mae(as.numeric(pred2$mean),moderna.test),
           mse(as.numeric(pred2$mean),moderna.test))
a1 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
moderna.train <- ts(moderna.train)
fit2 <- Arima(moderna.train,order=c(1,1,3),include.drift=TRUE)
moderna.sar <- Arima(moderna.train,order=c(1,1,3),seasonal=c(0,1,1))
autoplot(moderna.train,series='Data') +
  autolayer(meanf(moderna.train,150),PI=FALSE,series='Mean') +
  autolayer(naive(moderna.train,150),PI=FALSE,series='Naive') +
  autolayer(rwf(moderna.train,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  autolayer(forecast(moderna.sar,150),PI=FALSE, series='SAR') +
  xlab("Date") + ylab("Value") +
  ggtitle("modernaBenchmark Stock Forecast")
```

### vaccine 1
#### time series Plot
```{r}
vaccine.ts1 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Johnson&Johnson'),c(2,4)]))
vaccine.ts1 <- na.remove(vaccine.ts1)
acf(vaccine.ts1)
```
#### b ADF Test
```{r}
adf.test(vaccine.ts1)
```

```{r}
acf(log(vaccine.ts1))
adf.test(log(vaccine.ts1))
```

### 3
#### a Detrend Data
```{r}
v1.df<-diff(vaccine.ts1)
```

#### b Plot the differenced data and comment.
```{r}
fig1 <- autoplot(vaccine.ts1,series = 'Origin') + 
  autolayer(v1.df,series = 'Difference')
fig1 <- fig1 +labs(title = "Johnson&Johnson Origin and Difference",
                   x='Date',y='Value')
fig1
```

#### c ADF Test
```{r}
acf(v1.df)
adf.test(v1.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(v1.df)
pacf(v1.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(vaccine.ts1, order=c(1, 1, 1),include.drift = TRUE)
fit2 <- Arima(vaccine.ts1, order=c(1, 1, 2),include.drift = TRUE)
fit3 <- Arima(vaccine.ts1, order=c(4, 1, 1),include.drift = TRUE)
fit4 <- Arima(vaccine.ts1, order=c(4, 1, 2),include.drift = TRUE)
r4 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r4) <- c('Model','AIC','BIC','AICC')
r4[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r4[2,]<- c('ARIMA(1,1,2)',fit2$aic,fit2$bic,fit2$aicc)
r4[3,]<- c('ARIMA(4,1,1)',fit3$aic,fit3$bic,fit3$aicc)
r4[4,]<- c('ARIMA(4,1,2)',fit4$aic,fit4$bic,fit4$aicc)
r4 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
v1.train <- vaccine.ts1[1:floor(length(vaccine.ts1)*0.8)]
v1.test <- vaccine.ts1[(ceiling(length(vaccine.ts1)*0.8)+1):length(vaccine.ts1)]
v1.ar <- arima(v1.train, order=c(1,1,2))
summary(v1.ar)
```
$$  X_{t} =  0.4363X_{t-1}+0.6637X_{t-2}+w_{t}+0.1159w_{t-1}-0.6856w_{t-2}$$
### 6
#### full model diagnostics
```{r}
sarima(v1.test,1,1,2)
```

### 7 
#### auto.arima
```{r}
auto.arima(v1.train)
```

### 8
#### Forecast
```{r}
pred1 <- forecast(v1.ar,length(v1.test)+50)
autoplot(pred1) 
autoplot(vaccine.ts1)
```


### 9 
#### Benchmark
```{r}
f1 <- meanf(v1.train,length(v1.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(v1.train,length(v1.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(v1.train,length(v1.test)+50)
accuracy(f3)
checkresiduals(f3)
```

```{r message=FALSE, warning=FALSE}
library(Metrics)
a4 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a1) <- c('Method','MAE','MSE')
a4[1,] <- c('Mean',mae(as.numeric(f1$mean),v1.test),
           mse(as.numeric(f1$mean),v1.test))
a4[2,] <- c('Naive',mae(as.numeric(f2$mean),v1.test),
           mse(as.numeric(f2$mean),v1.test))
a4[3,] <- c('Drift',mae(as.numeric(f3$mean),v1.test),
           mse(as.numeric(f3$mean),v1.test))
a4[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),v1.test),
           mse(as.numeric(pred1$mean),v1.test))
a4 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
fit2 <- Arima(vaccine.ts1,order=c(1,1,3),include.drift=TRUE)
autoplot(vaccine.ts1,series='Data') +
  autolayer(meanf(vaccine.ts1,150),PI=FALSE,series='Mean') +
  autolayer(naive(vaccine.ts1,150),PI=FALSE,series='Naive') +
  autolayer(rwf(vaccine.ts1,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Benchmark Vaccine Forecast")
```

### vaccine 2
#### time series Plot
```{r}
vaccine.ts2 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Pfizer/BioNTech'),c(2,4)]))
vaccine.ts2 <- na.remove(vaccine.ts2)
acf(vaccine.ts2)
```
#### b ADF Test
```{r}
adf.test(vaccine.ts2)
```
```{r}
acf(log(vaccine.ts2))
adf.test(log(vaccine.ts2))
```

### 3
#### a Detrend Data
```{r}
v2.df<-diff(vaccine.ts2)
```

#### b Plot the differenced data and comment.
```{r}
fig1 <- autoplot(vaccine.ts2,series = 'Origin') + 
  autolayer(v2.df,series = 'Difference')
fig1 <- fig1 +labs(title = "Pfizer Vaccine and Difference",
                   x='Date',y='Value')
fig1
```

#### c ADF Test
```{r}
acf(v2.df)
adf.test(v2.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(v2.df)
pacf(v2.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(vaccine.ts2, order=c(1, 1, 1),include.drift = TRUE)
fit2 <- Arima(vaccine.ts2, order=c(1, 1, 2),include.drift = TRUE)
r5 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r5) <- c('Model','AIC','BIC','AICC')
r5[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r5[2,]<- c('ARIMA(1,1,2)',fit2$aic,fit2$bic,fit2$aicc)
r5 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
v2.train <- vaccine.ts2[1:floor(length(vaccine.ts2)*0.8)]
v2.test <- vaccine.ts2[(ceiling(length(vaccine.ts2)*0.8)+1):length(vaccine.ts2)]
v2.ar <- arima(v2.train, order=c(1,1,1))
summary(v2.ar)
```
$$  X_{t} =  1.1402X_{t-1}-0.1402X_{t-2}+w_{t}-0.7740w_{t-1}$$
### 6
#### full model diagnostics
```{r}
sarima(v2.train,1,1,1)
```

### 7 
#### auto.arima
```{r}
auto.arima(v2.train)
```

### 8
#### Forecast
```{r}
pred1 <- forecast(v2.ar,length(v2.test)+50)
autoplot(pred1) 
autoplot(vaccine.ts2)
```


### 9 
#### Benchmark
```{r}
f1 <- meanf(v2.train,length(v2.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(v2.train,length(v2.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(v2.train,length(v2.test)+50)
accuracy(f3)
checkresiduals(f3)
```

```{r message=FALSE, warning=FALSE}
library(Metrics)
a5 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a5) <- c('Method','MAE','MSE')
a5[1,] <- c('Mean',mae(as.numeric(f1$mean),v2.test),
           mse(as.numeric(f1$mean),v2.test))
a5[2,] <- c('Naive',mae(as.numeric(f2$mean),v2.test),
           mse(as.numeric(f2$mean),v2.test))
a5[3,] <- c('Drift',mae(as.numeric(f3$mean),v2.test),
           mse(as.numeric(f3$mean),v2.test))
a5[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),v2.test),
           mse(as.numeric(pred1$mean),v2.test))
a5 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
fit2 <- Arima(vaccine.ts2,order=c(1,1,3),include.drift=TRUE)
autoplot(vaccine.ts2,series='Data') +
  autolayer(meanf(vaccine.ts2,150),PI=FALSE,series='Mean') +
  autolayer(naive(vaccine.ts2,150),PI=FALSE,series='Naive') +
  autolayer(rwf(vaccine.ts2,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer Benchmark Vaccine Forecast")
```

### vaccine 3
#### time series Plot
```{r}
vaccine.ts3 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Moderna'),c(2,4)]))
vaccine.ts3 <- na.remove(vaccine.ts3)
acf(vaccine.ts3)
```
#### b ADF Test
```{r}
adf.test(vaccine.ts3)
```
### 3
#### a Detrend Data
```{r}
v3.df<-diff(vaccine.ts3)
```

#### b Plot the differenced data and comment.
```{r}
fig1 <- autoplot(vaccine.ts3,series = 'Origin') + 
  autolayer(v3.df,series = 'Difference')
fig1 <- fig1 +labs(title = "Moderna Origin and Difference",
                   x='Date',y='Value')
fig1
```

#### c ADF Test
```{r}
acf(v3.df)
adf.test(v3.df)
```

### 4
#### Plot the ACF and PACF plots 
```{r}
par(mfrow=c(1,2))
acf(v3.df)
pacf(v3.df)
```

### 5
#### check AIC and BIC

```{r}
fit1 <- Arima(vaccine.ts3, order=c(1, 1, 1),include.drift = TRUE)
fit2 <- Arima(vaccine.ts3, order=c(1, 1, 2),include.drift = TRUE)
fit3 <- Arima(vaccine.ts3, order=c(4, 1, 1),include.drift = TRUE)
fit4 <- Arima(vaccine.ts3, order=c(4, 1, 2),include.drift = TRUE)
r6 <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(r6) <- c('Model','AIC','BIC','AICC')
r6[1,]<- c('ARIMA(1,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r6[2,]<- c('ARIMA(1,1,2)',fit2$aic,fit2$bic,fit2$aicc)
r6[3,]<- c('ARIMA(4,1,1)',fit1$aic,fit1$bic,fit1$aicc)
r6[4,]<- c('ARIMA(4,1,2)',fit2$aic,fit2$bic,fit2$aicc)
r6 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
```

#### Train
```{r}
v3.train <- vaccine.ts3[1:floor(length(vaccine.ts3)*0.8)]
v3.test <- vaccine.ts3[(ceiling(length(vaccine.ts3)*0.8)+1):length(vaccine.ts3)]
v3.ar <- arima(v3.train, order=c(1,1,1))
summary(v3.ar)
```
$$  X_{t} =  1.3443X_{t-1}-0.3443X_{t-2}+w_{t}-0.8576w_{t-1}$$
### 6
#### full model diagnostics
```{r}
sarima(v3.train,1,1,1)
```

### 7 
#### auto.arima
```{r}
auto.arima(v3.train)
```

### 8
#### Forecast
```{r}
pred1 <- forecast(v3.ar,length(v3.test)+50)
autoplot(pred1) 
autoplot(vaccine.ts3)
```


### 9 
#### Benchmark
```{r}
f1 <- meanf(v3.train,length(v3.test)+50)
accuracy(f1)
checkresiduals(f1)

f2 <- naive(v3.train,length(v3.test)+50)
accuracy(f2)
checkresiduals(f2)

f3 <- rwf(v3.train,length(v3.test)+50)
accuracy(f3)
checkresiduals(f3)
```

```{r message=FALSE, warning=FALSE}
library(Metrics)
a6 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(a6) <- c('Method','MAE','MSE')
a6[1,] <- c('Mean',mae(as.numeric(f1$mean),v3.test),
           mse(as.numeric(f1$mean),v3.test))
a6[2,] <- c('Naive',mae(as.numeric(f2$mean),v3.test),
           mse(as.numeric(f2$mean),v3.test))
a6[3,] <- c('Drift',mae(as.numeric(f3$mean),v3.test),
           mse(as.numeric(f3$mean),v3.test))
a6[4,] <- c('ARIMA',mae(as.numeric(pred1$mean),v3.test),
           mse(as.numeric(pred1$mean),v3.test))
a6 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")
detach(package:Metrics)
```

```{r}
fit2 <- Arima(vaccine.ts3,order=c(1,1,3),include.drift=TRUE)
autoplot(vaccine.ts3,series='Data') +
  autolayer(meanf(vaccine.ts3,150),PI=FALSE,series='Mean') +
  autolayer(naive(vaccine.ts3,150),PI=FALSE,series='Naive') +
  autolayer(rwf(vaccine.ts3,150),PI=FALSE,series='Drift') +
  autolayer(forecast(fit2,150),PI=FALSE, series='AR') +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Benchmark Vaccine Forecast")
```
