---
title: "Exploratory Data Analysis"
author: "Lin Meng"
date: "2022/2/7"
output: html_document
---

```{r}
library(ggfortify)
library(ggplot2)
library(fpp)
library(stats)
library(tsbox)
library(gridExtra)
```

## Load Data
```{r}
covid <- read.csv("data/covid.csv")
johnson <- read.csv("data/Johnson.csv ")
moderna <- read.csv("data/Moderna.csv ")
pfizer <- read.csv("data/Pfizer.csv ")
stock <- read.csv("data/stock.csv")
vaccine <- read.csv("data/vaccine.csv")
```


## EDA

### johnson 

#### time series Plot
```{r}
johnson.ts <- ts_ts(ts_long(johnson[,c(1,5)]))
johnson.ts<-na.remove(johnson.ts)
fig1<-autoplot(johnson.ts,colour='#029386') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig1
```


#### Lag Plot
```{r}
gglagplot(johnson.ts, do.lines=FALSE,set.lags=1:9)
```

#### Decompose 
```{r}
decompose_johnson <- decompose(johnson.ts,'additive')
plot(decompose_johnson)
```

```{r}
decompose_johnson2 <- decompose(johnson.ts,'multiplicative')
plot(decompose_johnson2)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
par(mfrow=c(1,2))
acf(johnson.ts)
pacf(johnson.ts)
```

#### adf.test
```{r}
adf.test(johnson.ts)
```
#### Detrend Data
```{r}
johnson.df<-diff(log(johnson.ts))
adf.test(johnson.df)
```

```{r}
par(mfrow=c(1,2))
acf(johnson.ts)
acf(johnson.df)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(johnson.ts, series="Data") +
  autolayer(na.remove(ma(johnson.ts,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Moving Average Smoothing 3") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(johnson.ts, series="Data") +
  autolayer(na.remove(ma(johnson.ts,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Moving Average Smoothing 10") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma3
```
windows 20
```{r}
ma20 <- autoplot(johnson.ts, series="Data") +
  autolayer(na.remove(ma(johnson.ts,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### Moderna 
#### time series Plot
```{r}
moderna.ts <- ts_ts(ts_long(moderna[,c(1,5)]))
moderna.ts<-na.remove(moderna.ts)
fig2<-autoplot(moderna.ts,colour='#ff796c') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(moderna.ts, do.lines=FALSE,set.lags=1:9)
```

#### Decompose 
```{r}
decompose_moderna <- decompose(moderna.ts,'additive')
plot(decompose_moderna)
```

```{r}
decompose_moderna2 <- decompose(moderna.ts,'multiplicative')
plot(decompose_moderna2)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(moderna.ts)
pacf(moderna.ts)
```

#### adf.test
```{r}
adf.test(moderna.ts)
```

#### Detrend Data
```{r}
moderna.df<-diff(log(moderna.ts))
adf.test(moderna.df)
```

```{r}
par(mfrow=c(1,2))
acf(moderna.ts)
acf(moderna.df)
```
#### Moving Average
windows 3
```{r}
ma3 <- autoplot(moderna.ts, series="Data") +
  autolayer(na.remove(ma(moderna.ts,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(moderna.ts, series="Data") +
  autolayer(na.remove(ma(moderna.ts,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(moderna.ts, series="Data") +
  autolayer(na.remove(ma(moderna.ts,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### pfizer 
#### time series Plot
```{r}
pfizer.ts <- ts_ts(ts_long(pfizer[,c(1,5)]))
pfizer.ts<-na.remove(pfizer.ts)
fig2<-autoplot(pfizer.ts,colour='#e6daa6') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(pfizer.ts, do.lines=FALSE,set.lags=1:9)
```


#### Decompose 
```{r}
decompose_pfizer <- decompose(pfizer.ts,'additive')
plot(decompose_pfizer)
```

```{r}
decompose_pfizer2 <- decompose(pfizer.ts,'multiplicative')
plot(decompose_pfizer2)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(pfizer.ts)
pacf(pfizer.ts)
```

#### adf.test
```{r}
adf.test(pfizer.ts)
```

#### Detrend Data
```{r}
pfizer.df<-diff(log(pfizer.ts))
adf.test(pfizer.df)
```

```{r}
par(mfrow=c(1,2))
acf(pfizer.ts)
acf(pfizer.df)
```
#### Moving Average
windows 3
```{r}
ma3 <- autoplot(pfizer.ts, series="Data") +
  autolayer(na.remove(ma(pfizer.ts,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(pfizer.ts, series="Data") +
  autolayer(na.remove(ma(pfizer.ts,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(pfizer.ts, series="Data") +
  autolayer(na.remove(ma(pfizer.ts,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### stock 
#### time series Plot
```{r}
stock.ts <- ts_ts(ts_long(stock[,c(2,5)]))
stock.ts<-na.remove(stock.ts)
fig2<-autoplot(stock.ts,colour='#703be7') + labs(x="Date", y="Closed", title="Stock Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(stock.ts, do.lines=FALSE,set.lags=1:9)
```


#### Decompose 
```{r}
decompose_stock <- decompose(stock.ts,'additive')
plot(decompose_stock)
```

```{r}
decompose_stock2 <- decompose(stock.ts,'multiplicative')
plot(decompose_stock2)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(stock.ts)
pacf(stock.ts)
```

#### adf.test
```{r}
adf.test(stock.ts)
```

#### Detrend Data
```{r}
stock.df<-diff(log(stock.ts))
adf.test(stock.df)
```

```{r}
par(mfrow=c(1,2))
acf(stock.ts)
acf(stock.df)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(stock.ts, series="Data") +
  autolayer(na.remove(ma(stock.ts,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Stock Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(stock.ts, series="Data") +
  autolayer(na.remove(ma(stock.ts,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Stock Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(stock.ts, series="Data") +
  autolayer(na.remove(ma(stock.ts,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Stock Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### covid 1 
#### time series Plot
```{r}
covid.ts1 <- ts_ts(ts_long(covid[,c(1,3)]))
covid.ts1<-na.remove(covid.ts1)
fig2<-autoplot(covid.ts1,colour='#0652ff') + labs(x="Date", y="New Cases", title="covid Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(covid.ts1, do.lines=FALSE,set.lags=1:9)
```


#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(covid.ts1)
pacf(covid.ts1)
```

#### adf.test
```{r}
adf.test(covid.ts1)
```

#### Detrend Data
```{r}
covid.df1<-diff(covid.ts1)
adf.test(covid.df1)
```

```{r}
par(mfrow=c(1,2))
acf(covid.ts1)
acf(covid.df1)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(covid.ts1, series="Data") +
  autolayer(na.remove(ma(covid.ts1,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid New Caeses Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(covid.ts1, series="Data") +
  autolayer(na.remove(ma(covid.ts1,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid New Caeses Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(covid.ts1, series="Data") +
  autolayer(na.remove(ma(covid.ts1,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid New Caeses Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### covid 2
#### time series Plot
```{r}
covid.ts2 <- ts_ts(ts_long(covid[,c(1,4)]))
covid.ts2<-na.remove(covid.ts2)
fig2<-autoplot(covid.ts2,colour='#ffa756') + labs(x="Date", y="New Cases", title="covid Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(covid.ts2, do.lines=FALSE,set.lags=1:9)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(covid.ts2)
pacf(covid.ts2)
```

#### adf.test
```{r}
adf.test(covid.ts2)
```

#### Detrend Data
```{r}
covid.df2<-diff(covid.ts2)
adf.test(covid.df2)
```

```{r}
par(mfrow=c(1,2))
acf(covid.ts2)
acf(covid.df2)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(covid.ts2, series="Data") +
  autolayer(na.remove(ma(covid.ts2,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid Deaths Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(covid.ts2, series="Data") +
  autolayer(na.remove(ma(covid.ts2,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid Deaths Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(covid.ts2, series="Data") +
  autolayer(na.remove(ma(covid.ts2,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Covid Deaths Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### vaccine 1
#### time series Plot
```{r}
vaccine.ts1 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Moderna'),c(2,4)]))
vaccine.ts1<-na.remove(vaccine.ts1)
fig2<-autoplot(vaccine.ts1,colour='#cb0162') + labs(x="Date", y="New Cases", title="vaccine Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(vaccine.ts1, do.lines=FALSE,set.lags=1:9)
```

#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(vaccine.ts1)
pacf(vaccine.ts1)
```

#### adf.test
```{r}
adf.test(vaccine.ts1)
```

#### Detrend Data
```{r}
vaccine.df1<-diff(vaccine.ts1)
adf.test(vaccine.df1)
```

```{r}
par(mfrow=c(1,2))
acf(vaccine.ts1)
acf(vaccine.df1)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(vaccine.ts1, series="Data") +
  autolayer(na.remove(ma(vaccine.ts1,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(vaccine.ts1, series="Data") +
  autolayer(na.remove(ma(vaccine.ts1,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(vaccine.ts1, series="Data") +
  autolayer(na.remove(ma(vaccine.ts1,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Moderna Vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### vaccine 2
#### time series Plot
```{r}
vaccine.ts2 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Johnson&Johnson'),c(2,4)]))
vaccine.ts2<-na.remove(vaccine.ts2)
fig2<-autoplot(vaccine.ts2,colour='#b00149') + labs(x="Date", y="New Cases", title="vaccine Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(vaccine.ts2, do.lines=FALSE,set.lags=1:9)
```



#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(vaccine.ts2)
pacf(vaccine.ts2)
```

#### adf.test
```{r}
adf.test(vaccine.ts2)
```

#### Detrend Data
```{r}
vaccine.df2<-diff(vaccine.ts2)
adf.test(vaccine.df2)
```

```{r}
par(mfrow=c(1,2))
acf(vaccine.ts2)
acf(vaccine.df2)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(vaccine.ts2, series="Data") +
  autolayer(na.remove(ma(vaccine.ts2,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(vaccine.ts2, series="Data") +
  autolayer(na.remove(ma(vaccine.ts2,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(vaccine.ts2, series="Data") +
  autolayer(na.remove(ma(vaccine.ts2,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Johnson&Johnson Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```

### vaccine 3
#### time series Plot
```{r}
vaccine.ts3 <- ts_ts(ts_long(vaccine[which(vaccine$vaccine=='Pfizer/BioNTech'),c(2,4)]))
vaccine.ts3<-na.remove(vaccine.ts3)
fig2<-autoplot(vaccine.ts3,colour='#b66a50') + labs(x="Date", y="New Cases", title="vaccine Closed Value By Date")
fig2
```

#### Lag Plot
```{r}
gglagplot(vaccine.ts3, do.lines=FALSE,set.lags=1:9)
```


#### ACF&PACF
```{r}
par(mfrow=c(1,2))
acf(vaccine.ts3)
pacf(vaccine.ts3)
```

#### adf.test
```{r}
adf.test(vaccine.ts3)
```

#### Detrend Data
```{r}
vaccine.df3<-diff(vaccine.ts3)
adf.test(vaccine.df3)
```

```{r}
par(mfrow=c(1,2))
acf(vaccine.ts3)
acf(vaccine.df3)
```

#### Moving Average
windows 3
```{r}
ma3 <- autoplot(vaccine.ts3, series="Data") +
  autolayer(na.remove(ma(vaccine.ts3,3)), series="3-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
  breaks=c("Data","3-MA"))+ theme_bw()
ma3
```
windows 10
```{r}
ma10 <- autoplot(vaccine.ts3, series="Data") +
  autolayer(na.remove(ma(vaccine.ts3,10)), series="10-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="grey50","10-MA"='blue'),
  breaks=c("Data","10-MA"))+ theme_bw()
ma20
```
windows 20
```{r}
ma20 <- autoplot(vaccine.ts3, series="Data") +
  autolayer(na.remove(ma(vaccine.ts3,20)), series="20-MA") +
  xlab("Date") + ylab("Value") +
  ggtitle("Pfizer vaccine Moving Average Smoothing 20") +
  scale_colour_manual(values=c("Data"="#acaaa6","20-MA"='purple'),
  breaks=c("Data","20-MA"))+ theme_bw()
ma20
```